### 2019-11-20 ic-journal/part2.tex
Cumulative fixes with highlights

Based on results about ``bisimulation up to $\sim$'', we have first proved the following
 lemma. It states  that if $X$ is weakly guarded
in $E$, then the ``first move'' of $E$ is independent of the agent
substituted for $X$:
\begin{lemma}[\texttt{STRONG_UNIQUE_SOLUTION_LEMMA} (Lemma 3.13 of~\cite{Mil89}, \univariate version)]
\label{lem:313}
If the variable $X$ is weakly guarded in $E$, and
$E\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$E'\{P/X\}$ (for some expression $E'$), and moreover, for any $Q$,
$E\{Q/X\}\overset{\alpha}{\rightarrow} E'\{Q/X\}$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \HOLBoundVar{P\sp{\prime}}. \HOLFreeVar{E} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sp{\prime}}. \HOLConst{CONTEXT} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{E\sp{\prime}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{E} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{E\sp{\prime}} \HOLBoundVar{Q}
\end{alltt}
\end{lemma}

@@

Milner's book~\cite{Mil89} contains only two ``unique solutions of
equations'' theorems, one for strong equivalence, the other for
observational congruence.   There is however a
version  of the result for weak bisimilarity that shares a large portion of proof
steps with that case for observational congruence.  
As  weak bisimilarity 
 is not \hl{a} congruence, we have to be more restrictive on the syntax for
 the equation, using strong guardedness.
% to make correct statement of this
% theorem, we must slightly modify the concept of sequential expresses
% with further restrictions: no direct sums (or only guarded sums).
% The ``unique solution of equations'' theorem for weak equivalence
% requires more restrictions on the expression, namely, .
% Now we're ready to state and prove the following lemma:
\begin{lemma}[\texttt{WEAK_UNIQUE_SOLUTION_LEMMA}]
If the variable $X$ is guarded and sequential in $G$, and
$G\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$H\{P/X\}$, for some expression $H$, and for any $Q$, we have
$G\{Q/X\}\overset{\alpha}{\rightarrow} H\{Q/X\}$. Moreover $H$ is
sequential, and if $\alpha = \tau$ then $H$ is also guarded.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{G} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLFreeVar{G} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \HOLBoundVar{P\sp{\prime}}.
       \HOLFreeVar{G} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{H}. \HOLConst{GSEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLBoundVar{a} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} \HOLBoundVar{H}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{H} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{G} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{H} \HOLBoundVar{Q}
\end{alltt}
\end{lemma}

@@

\begin{lemma}[\texttt{OBS_CONGR_BY_WEAK_BISIM}]
% To prove two processes $E$ and $E'$  observationally congruent, it is
% sufficient enough to construct a bisimulation with additional observational
% transition properties:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WEAK_BISIM} \HOLFreeVar{Wbsm} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{E\sp{\prime}}.
       \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{u}.
            \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E\sb{\mathrm{1}}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sb{\mathrm{2}}}. \HOLBoundVar{E\sp{\prime}} \HOLTokenWeakTransBegin\HOLBoundVar{u}\HOLTokenWeakTransEnd \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Wbsm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
            \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E\sb{\mathrm{2}}}. \HOLBoundVar{E\sp{\prime}} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{E\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sb{\mathrm{1}}}. \HOLBoundVar{E} \HOLTokenWeakTransBegin\HOLBoundVar{u}\HOLTokenWeakTransEnd \HOLBoundVar{E\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Wbsm} \HOLBoundVar{E\sb{\mathrm{1}}} \HOLBoundVar{E\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLBoundVar{E} \HOLSymConst{\HOLTokenObsCongr} \HOLBoundVar{E\sp{\prime}}
\end{alltt}
\end{lemma}

This lemma plays a key role in the
corresponding `unique solutions' theorem, in connection with the
following result. 
In this part, we have not 
used 
 ``bisimulation up-to'' techniques. Using
 the  above \texttt{OBS_CONGR_BY_WEAK_BISIM},
 we  directly construct the appropriate bisimulation relations.
\begin{lemma}[\texttt{OBS_UNIQUE_SOLUTION_LEMMA}]
If the variable $X$ is guarded and sequential in $G$, and
$G\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$H\{P/X\}$, for some  $H$, and for any $Q$, we also have 
$G\{Q/X\}\overset{\alpha}{\rightarrow} H\{Q/X\}$. Moreover $H$ is
sequential, and if $\alpha = \tau$ then $H$ is also guarded.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{G} \HOLSymConst{\HOLTokenConj{}} \HOLConst{SEQ} \HOLFreeVar{G} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \HOLBoundVar{P\sp{\prime}}.
       \HOLFreeVar{G} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{H}. \HOLConst{SEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLBoundVar{a} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} \HOLBoundVar{H}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{H} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{G} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{H} \HOLBoundVar{Q}
\end{alltt}
\end{lemma}

@@

A trace is represented by the initial and final processes, plus
a list of actions  so performed.
For this, we first 
 define the concept of label-accumulated reflexive transitive closure
 (\texttt{LRTC}).
Given a labeled transition relation \texttt{R} on CCS, \texttt{LRTC R} is
a label-accumulated relation representing the trace of transitions:
\begin{alltt}
   \HOLConst{LRTC} \HOLFreeVar{R} \HOLFreeVar{a} \HOLFreeVar{l} \HOLFreeVar{b} \HOLSymConst{\HOLTokenDefEquality{}}
     \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P}.
         \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x}. \HOLBoundVar{P} \HOLBoundVar{x} \ensuremath{[}\ensuremath{]} \HOLBoundVar{x}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
         \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x} \HOLBoundVar{h} \HOLBoundVar{y} \HOLBoundVar{t} \HOLBoundVar{z}. \HOLFreeVar{R} \HOLBoundVar{x} \HOLBoundVar{h} \HOLBoundVar{y} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P} \HOLBoundVar{y} \HOLBoundVar{t} \HOLBoundVar{z} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLBoundVar{x} \ensuremath{(}\HOLBoundVar{h}\HOLSymConst{::}\HOLBoundVar{t}\ensuremath{)} \HOLBoundVar{z}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
         \HOLBoundVar{P} \HOLFreeVar{a} \HOLFreeVar{l} \HOLFreeVar{b}\hfill{[LRTC_DEF]}
\end{alltt}
The trace relation for CCS can be then obtained
 by calling \texttt{LRTC} on the (strong, or single-step) labeled transition
 relation \texttt{TRANS} ($\overset{\mu}{\rightarrow}$) defined by SOS rules:
\begin{alltt}
   \HOLConst{TRACE} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{LRTC} \HOLConst{TRANS}\hfill{[TRACE_def]}
\end{alltt}

If the list of actions is empty, that means that there is no transition and hence,
if there is at most one visible action (i.e., a label) in the list of actions,
then the trace is also a weak transition. Here
we have to distinguish between two cases: no label and unique label (in
the list of actions). The definition of ``no
label'' in an action list is easy (here \texttt{MEM} tests if a given element is a member of a list):
\begin{alltt}
   \HOLConst{NO_LABEL} \HOLFreeVar{L} \HOLSymConst{\HOLTokenDefEquality{}} \HOLSymConst{\HOLTokenNeg{}}\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{l}. \HOLConst{MEM} \ensuremath{(}\HOLConst{label} \HOLBoundVar{l}\ensuremath{)} \HOLFreeVar{L}\hfill{[NO_LABEL_def]}
\end{alltt}

The definition of ``unique label'' can be done in many \hl{ways. The}
following definition (a suggestion from Robert Beers)
avoids any counting or filtering in the list.
It says that a label is unique in a list of actions if and only if there is no
label in the rest of list:
\begin{alltt}
   \HOLConst{UNIQUE_LABEL} \HOLFreeVar{u} \HOLFreeVar{L} \HOLSymConst{\HOLTokenDefEquality{}}
     \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{L\sb{\mathrm{1}}} \HOLBoundVar{L\sb{\mathrm{2}}}. \HOLBoundVar{L\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenDoublePlus} \ensuremath{[}\HOLFreeVar{u}\ensuremath{]} \HOLSymConst{\HOLTokenDoublePlus} \HOLBoundVar{L\sb{\mathrm{2}}} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{L} \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \HOLBoundVar{L\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \HOLBoundVar{L\sb{\mathrm{2}}}\hfill{[UNIQUE_LABEL_def]}
\end{alltt}

The final relationship between traces and weak transitions is stated
and proved in the following theorem
(where the  variable $acts$ stands for
a list of actions); 
it says, a weak transition $P\overset{u}{\Rightarrow}P'$ is also a
trace $P\overset{acts}{\longrightarrow}P'$ with a
 non-empty action list $acts$, in which either there is no label (for $u = \tau$), or 
$u$ is the unique label (for $u \neq \tau$):
%\begin{lemma}
% A weak transition $P\overset{u}{\Rightarrow}P'$ is a just trace with non
% empty action list: 1) without any visible label, if $u = \tau$, or 2)
% $u$ is the unique label in the list, if $u \neq \tau$.
\begin{alltt}
\HOLTokenTurnstile{} \HOLFreeVar{P} \HOLTokenWeakTransBegin\HOLFreeVar{u}\HOLTokenWeakTransEnd \HOLFreeVar{P\sp{\prime}} \HOLSymConst{\HOLTokenEquiv{}}
   \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{acts}.
       \HOLConst{TRACE} \HOLFreeVar{P} \HOLBoundVar{acts} \HOLFreeVar{P\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenNeg{}}\HOLConst{NULL} \HOLBoundVar{acts} \HOLSymConst{\HOLTokenConj{}}
       \HOLKeyword{if} \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLKeyword{then} \HOLConst{NO_LABEL} \HOLBoundVar{acts} \HOLKeyword{else} \HOLConst{UNIQUE_LABEL} \HOLFreeVar{u} \HOLBoundVar{acts}\hfill{[WEAK_TRANS_AND_TRACE]}
\end{alltt}
%\end{lemma}

>>>>

Based on results about ``bisimulation up to $\sim$'', we have first proved the following
 lemma. It states  that if $X$ is weakly guarded
in $E$, then the ``first move'' of $E$ is independent of the agent
substituted for $X$:
\begin{lemma}[\texttt{STRONG_UNIQUE_SOLUTION_LEMMA} (Lemma 3.13 of~\cite{Mil89}, \univariate version)]
\label{lem:313}
If the variable $X$ is weakly guarded in $E$, and
$E\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$E'\{P/X\}$ (for some expression $E'$), and moreover, for any $Q$,
$E\{Q/X\}\overset{\alpha}{\rightarrow} E'\{Q/X\}$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WG} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \ensuremath{\HOLBoundVar{P}\sp{\prime}}. \HOLFreeVar{E} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLConst{CONTEXT} \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\ensuremath{=}} \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{E} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLBoundVar{Q}
\end{alltt}
\end{lemma}

@@

Milner's book~\cite{Mil89} contains only two ``unique solutions of
equations'' theorems, one for strong equivalence, the other for
observational congruence.   There is however a
version  of the result for weak bisimilarity that shares a large portion of proof
steps with that case for observational congruence.  
As  weak bisimilarity 
 is not \hl{a} congruence, we have to be more restrictive on the syntax for
 the equation, using strong guardedness.
% to make correct statement of this
% theorem, we must slightly modify the concept of sequential expresses
% with further restrictions: no direct sums (or only guarded sums).
% The ``unique solution of equations'' theorem for weak equivalence
% requires more restrictions on the expression, namely, .
% Now we're ready to state and prove the following lemma:
\begin{lemma}[\texttt{WEAK_UNIQUE_SOLUTION_LEMMA}]
If the variable $X$ is guarded and sequential in $G$, and
$G\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$H\{P/X\}$, for some expression $H$, and for any $Q$, we have
$G\{Q/X\}\overset{\alpha}{\rightarrow} H\{Q/X\}$. Moreover $H$ is
sequential, and if $\alpha = \tau$ then $H$ is also guarded.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{G} \HOLSymConst{\HOLTokenConj{}} \HOLConst{GSEQ} \HOLFreeVar{G} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \ensuremath{\HOLBoundVar{P}\sp{\prime}}.
       \HOLFreeVar{G} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{H}. \HOLConst{GSEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLBoundVar{a} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} \HOLBoundVar{H}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{H} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{G} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{H} \HOLBoundVar{Q}
\end{alltt}
\end{lemma}

@@

\begin{lemma}[\texttt{OBS_CONGR_BY_WEAK_BISIM}]
% To prove two processes $E$ and $E'$  observationally congruent, it is
% sufficient enough to construct a bisimulation with additional observational
% transition properties:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WEAK_BISIM} \HOLFreeVar{Wbsm} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \ensuremath{\HOLBoundVar{E}\sp{\prime}}.
       \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{u}.
            \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}. \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLTokenWeakTransBegin\HOLBoundVar{u}\HOLTokenWeakTransEnd \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Wbsm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
            \HOLSymConst{\HOLTokenForall{}}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}. \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenExists{}}\ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}}. \HOLBoundVar{E} \HOLTokenWeakTransBegin\HOLBoundVar{u}\HOLTokenWeakTransEnd \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLFreeVar{Wbsm} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{E}\sb{\mathrm{2}}}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
       \HOLBoundVar{E} \HOLSymConst{\HOLTokenObsCongr} \ensuremath{\HOLBoundVar{E}\sp{\prime}}
\end{alltt}
\end{lemma}

This lemma plays a key role in the
corresponding `unique solutions' theorem, in connection with the
following result. 
In this part, we have not 
used 
 ``bisimulation up-to'' techniques. Using
 the  above \texttt{OBS_CONGR_BY_WEAK_BISIM},
 we  directly construct the appropriate bisimulation relations.
\begin{lemma}[\texttt{OBS_UNIQUE_SOLUTION_LEMMA}]
If the variable $X$ is guarded and sequential in $G$, and
$G\{P/X\}\overset{\alpha}{\rightarrow} P'$, then $P'$ takes the form
$H\{P/X\}$, for some  $H$, and for any $Q$, we also have 
$G\{Q/X\}\overset{\alpha}{\rightarrow} H\{Q/X\}$. Moreover $H$ is
sequential, and if $\alpha = \tau$ then $H$ is also guarded.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{SG} \HOLFreeVar{G} \HOLSymConst{\HOLTokenConj{}} \HOLConst{SEQ} \HOLFreeVar{G} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P} \HOLBoundVar{a} \ensuremath{\HOLBoundVar{P}\sp{\prime}}.
       \HOLFreeVar{G} \HOLBoundVar{P} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{H}. \HOLConst{SEQ} \HOLBoundVar{H} \HOLSymConst{\HOLTokenConj{}} \ensuremath{(}\HOLBoundVar{a} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{SG} \HOLBoundVar{H}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \ensuremath{\HOLBoundVar{P}\sp{\prime}} \HOLSymConst{\ensuremath{=}} \HOLBoundVar{H} \HOLBoundVar{P} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Q}. \HOLFreeVar{G} \HOLBoundVar{Q} \HOLTokenTransBegin\HOLBoundVar{a}\HOLTokenTransEnd \HOLBoundVar{H} \HOLBoundVar{Q}
\end{alltt}
\end{lemma}

@@

A trace is represented by the initial and final processes, plus
a list of actions  so performed.
For this, we first 
 define the concept of label-accumulated reflexive transitive closure
 (\texttt{LRTC}).
Given a labeled transition relation \texttt{R} on CCS, \texttt{LRTC R} is
a label-accumulated relation representing the trace of transitions:
\begin{alltt}
   \HOLConst{LRTC} \HOLFreeVar{R} \HOLFreeVar{a} \HOLFreeVar{l} \HOLFreeVar{b} \HOLTokenDefEquality{}
     \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{P}.
         \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x}. \HOLBoundVar{P} \HOLBoundVar{x} \ensuremath{[}\ensuremath{]} \HOLBoundVar{x}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
         \ensuremath{(}\HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{x} \HOLBoundVar{h} \HOLBoundVar{y} \HOLBoundVar{t} \HOLBoundVar{z}. \HOLFreeVar{R} \HOLBoundVar{x} \HOLBoundVar{h} \HOLBoundVar{y} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P} \HOLBoundVar{y} \HOLBoundVar{t} \HOLBoundVar{z} \HOLSymConst{\HOLTokenImp{}} \HOLBoundVar{P} \HOLBoundVar{x} \ensuremath{(}\HOLBoundVar{h}\HOLSymConst{::}\HOLBoundVar{t}\ensuremath{)} \HOLBoundVar{z}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
         \HOLBoundVar{P} \HOLFreeVar{a} \HOLFreeVar{l} \HOLFreeVar{b}\hfill{[LRTC_DEF]}
\end{alltt}
The trace relation for CCS can be then obtained
 by calling \texttt{LRTC} on the (strong, or single-step) labeled transition
 relation \texttt{TRANS} ($\overset{\mu}{\rightarrow}$) defined by SOS rules:
\begin{alltt}
   \HOLConst{TRACE} \HOLTokenDefEquality{} \HOLConst{LRTC} \HOLConst{TRANS}\hfill{[TRACE_def]}
\end{alltt}

If the list of actions is empty, that means that there is no transition and hence,
if there is at most one visible action (i.e., a label) in the list of actions,
then the trace is also a weak transition. Here
we have to distinguish between two cases: no label and unique label (in
the list of actions). The definition of ``no
label'' in an action list is easy (here \texttt{MEM} tests if a given element is a member of a list):
\begin{alltt}
   \HOLConst{NO_LABEL} \HOLFreeVar{L} \HOLTokenDefEquality{} \HOLSymConst{\HOLTokenNeg{}}\HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{l}. \HOLConst{MEM} \ensuremath{(}\HOLConst{label} \HOLBoundVar{l}\ensuremath{)} \HOLFreeVar{L}\hfill{[NO_LABEL_def]}
\end{alltt}

The definition of ``unique label'' can be done in many \hl{ways. The}
following definition (a suggestion from Robert Beers)
avoids any counting or filtering in the list.
It says that a label is unique in a list of actions if and only if there is no
label in the rest of list:
\begin{alltt}
   \HOLConst{UNIQUE_LABEL} \HOLFreeVar{u} \HOLFreeVar{L} \HOLTokenDefEquality{}
     \HOLSymConst{\HOLTokenExists{}}\ensuremath{\HOLBoundVar{L}\sb{\mathrm{1}}} \ensuremath{\HOLBoundVar{L}\sb{\mathrm{2}}}. \ensuremath{\HOLBoundVar{L}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenDoublePlus} \ensuremath{[}\HOLFreeVar{u}\ensuremath{]} \HOLSymConst{\HOLTokenDoublePlus} \ensuremath{\HOLBoundVar{L}\sb{\mathrm{2}}} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{L} \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \ensuremath{\HOLBoundVar{L}\sb{\mathrm{1}}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{NO_LABEL} \ensuremath{\HOLBoundVar{L}\sb{\mathrm{2}}}\hfill{[UNIQUE_LABEL_def]}
\end{alltt}

The final relationship between traces and weak transitions is stated
and proved in the following theorem
(where the  variable $acts$ stands for
a list of actions); 
it says, a weak transition $P\overset{u}{\Rightarrow}P'$ is also a
trace $P\overset{acts}{\longrightarrow}P'$ with a
 non-empty action list $acts$, in which either there is no label (for $u = \tau$), or 
$u$ is the unique label (for $u \neq \tau$):
%\begin{lemma}
% A weak transition $P\overset{u}{\Rightarrow}P'$ is a just trace with non
% empty action list: 1) without any visible label, if $u = \tau$, or 2)
% $u$ is the unique label in the list, if $u \neq \tau$.
\begin{alltt}
\HOLTokenTurnstile{} \HOLFreeVar{P} \HOLTokenWeakTransBegin\HOLFreeVar{u}\HOLTokenWeakTransEnd \ensuremath{\HOLFreeVar{P}\sp{\prime}} \HOLSymConst{\HOLTokenEquiv{}}
   \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{acts}.
       \HOLConst{TRACE} \HOLFreeVar{P} \HOLBoundVar{acts} \ensuremath{\HOLFreeVar{P}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLSymConst{\HOLTokenNeg{}}\HOLConst{NULL} \HOLBoundVar{acts} \HOLSymConst{\HOLTokenConj{}}
       \HOLKeyword{if} \HOLFreeVar{u} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\ensuremath{\tau}} \HOLKeyword{then} \HOLConst{NO_LABEL} \HOLBoundVar{acts} \HOLKeyword{else} \HOLConst{UNIQUE_LABEL} \HOLFreeVar{u} \HOLBoundVar{acts}\hfill{[WEAK_TRANS_AND_TRACE]}
\end{alltt}
%\end{lemma}
